Tulai

### 2023-09-19 paper/paper_tulai.qmd
PCA-Color points

@@

```{r}
# Create the plot: WABI_Level_logscale
# here I do not consider those levels with a value 0 of retouch frequency. 
filtered_data <- sum_depth_sheet %>% 
  filter(lithic_dens > 0 & retouch_freq > 0)

@@

#PCA_level: for PCA Bicho and Cascalheira considered these variables: estimated area (as we calculated based on levels, I consider the thickness; is that OK?), core frequency, blank frequency, chip frequency: artefacts smaller than 1 cm (as we work on a neolithic asssemblage, we have so many small artefacts; so, I do not consider it), feature frequency (we have some stone alignments, but we do not know the exact level, we know the area of those features), retouch frequency, tool diversity (diversity of tool types within each assemblage, calculated using Menhinick’s index); 
##core frequency
##blank frequency
##retouch frequency
##tool diversity
##thickness of deposit (???)
##lithic density

@@

#save contribution plots into one objects
p1 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 1, 
                   top = 10)
p2 <- fviz_contrib(res.pca, 
                   choice = "var",
                   axes = 2, 
                   top = 10)
p3 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 3, 
                   top = 10)
p4 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 4, 
                   top = 10)

#combine individual plots into one plot 
library(gridExtra)
grid.arrange(p1, p2, p3, p4, ncol = 2)

@@

# Compute average contribution for each dimension
avg_contrib_dim1 <- mean(contributions[,1], na.rm = TRUE)
avg_contrib_dim2 <- mean(contributions[,2], na.rm = TRUE)
avg_contrib_dim3 <- mean(contributions[,3], na.rm = TRUE)
avg_contrib_dim4 <- mean(contributions[,4], na.rm = TRUE)

# Find variables that contribute more than average to each dimension
high_contrib_dim1 <- contributions[contributions[,1] > avg_contrib_dim1, 1]
high_contrib_dim2 <- contributions[contributions[,2] > avg_contrib_dim2, 2]
high_contrib_dim3 <- contributions[contributions[,3] > avg_contrib_dim3, 3]
high_contrib_dim4 <- contributions[contributions[,4] > avg_contrib_dim4, 4]

# Sort them in descending order
sorted_high_contrib_dim1 <- sort(high_contrib_dim1, decreasing = TRUE)
sorted_high_contrib_dim2 <- sort(high_contrib_dim2, decreasing = TRUE)
sorted_high_contrib_dim3 <- sort(high_contrib_dim3, decreasing = TRUE)
sorted_high_contrib_dim4 <- sort(high_contrib_dim4, decreasing = TRUE)

@@

@@

```{r}
# Visualize biplot:1
fviz_pca_biplot(res.pca,
                axes = c(1, 2),
                labelsize = 2
                )
rownames(res.pca$ind$coord) <- sum_depth_sheet$label

>>>>

summary_table_area_filtered <- summary_table_area %>%
  filter(area %in% c("ta", "d1", "tp1"))

library(kableExtra)

kable(summary_table_area_filtered, 
      caption = "Table: Summary by Area for 'ta', 'd1', and 'tp1'", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

@@

```{r}
library(ggpmisc)

# Create the plot: WABI_Level_logscale
# here I do not consider those levels with a value 0 of retouch frequency. 
filtered_data <- sum_depth_sheet %>% 
  filter(lithic_dens > 0 & retouch_freq > 0)

@@

#PCA_level: for PCA Bicho and Cascalheira considered these variables: estimated area (as we calculated based on levels, should I consider the thickness?), core frequency, blank frequency, chip frequency: artefacts smaller than 1 cm (as we work on a neolithic asssemblage, we have so many small artefacts; so, I do not consider it), feature frequency (we have some stone alignments, but we do not know the exact level, we know the area of those features), retouch frequency, tool diversity (diversity of tool types within each assemblage, calculated using Menhinick’s index); 
  ##core frequency
  ##blank frequency
  ##retouch frequency
  ##tool diversity
  ##thickness of deposit (???)
  ##lithic density

@@

#save contribution plots into one objects
p1 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 1, 
                   top = 10)
p2 <- fviz_contrib(res.pca, 
                   choice = "var",
                   axes = 2, 
                   top = 10)
p3 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 3, 
                   top = 10)
p4 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 4, 
                   top = 10)
p5 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 5, 
                   top = 10)

#combine individual plots into one plot 
library(gridExtra)
grid.arrange(p1, p2, p3, p4, p5, ncol = 2)

@@

# Compute average contribution for each dimension
avg_contrib_dim1 <- mean(contributions[,1], na.rm = TRUE)
avg_contrib_dim2 <- mean(contributions[,2], na.rm = TRUE)
avg_contrib_dim3 <- mean(contributions[,3], na.rm = TRUE)
avg_contrib_dim4 <- mean(contributions[,4], na.rm = TRUE)
avg_contrib_dim5 <- mean(contributions[,5], na.rm = TRUE)

# Find variables that contribute more than average to each dimension
high_contrib_dim1 <- contributions[contributions[,1] > avg_contrib_dim1, 1]
high_contrib_dim2 <- contributions[contributions[,2] > avg_contrib_dim2, 2]
high_contrib_dim3 <- contributions[contributions[,3] > avg_contrib_dim3, 3]
high_contrib_dim4 <- contributions[contributions[,4] > avg_contrib_dim4, 4]
high_contrib_dim5 <- contributions[contributions[,5] > avg_contrib_dim4, 5]

# Sort them in descending order
sorted_high_contrib_dim1 <- sort(high_contrib_dim1, decreasing = TRUE)
sorted_high_contrib_dim2 <- sort(high_contrib_dim2, decreasing = TRUE)
sorted_high_contrib_dim3 <- sort(high_contrib_dim3, decreasing = TRUE)
sorted_high_contrib_dim4 <- sort(high_contrib_dim4, decreasing = TRUE)
sorted_high_contrib_dim5 <- sort(high_contrib_dim5, decreasing = TRUE)

@@

print("Significant contributions to Dim5:")
print(sorted_high_contrib_dim5)

@@

```{r}
# Visualize biplot:1
fviz_pca_biplot(res.pca,
                axes = c(1, 2),
                labelsize = 2
                )

rownames(res.pca$ind$coord) <- sum_depth_sheet$label

```{r}
library(ggrepel)

# Generate the base biplot
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               labelsize = 2
                              )

# Modify row names of coordinates
rownames(res.pca$ind$coord) <- sum_depth_sheet$label

# Add geom_text_repel() to the base plot
biplot_modified <- biplot_base +
                   geom_text_repel(data = as.data.frame(res.pca$ind$coord),
                                   aes(x = Dim.1, 
                                       y = Dim.2, 
                                       label = rownames(res.pca$ind$coord)
                                       ),
                                   box.padding = 0.5,
                                   point.padding = 0.5,
                                   size = 2
                                  )

# Display the plot
biplot_modified

```{r}
# Base biplot without points
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               geom = "none",  
                               labelsize = 2
                               )

# Add points and text labels manually
biplot_modified <- biplot_base +
  geom_point(data = coord_df, aes(x = Dim.1, y = Dim.2, color = group), size = 1) +
  geom_text_repel(data = coord_df,
                  aes(x = Dim.1,
                      y = Dim.2,
                      label = label),
                  box.padding = 0.5,
                  point.padding = 0.5,
                  size = 2)

# Display the plot
biplot_modified
